---
title: "R Survey Cheat Sheet"
author: "David J. Barney^[davidjbarney@gmail.com]"
date: "1/26/2019"
output:
  pdf_document: default
  html_document: default
---


# Reading & Exploring Data


To read in your data:

```{r eval=FALSE}
# Comma delimited format
dat <- read.csv("~/your/file/path")
# Tab delimited format
dat <- read.table("~/your/file/path")
# Stata (.dta) format
library(haven)
dat <- read_dta("~/your/file/path")
```

To view observations or summaries of a variable:

```{r eval=FALSE}
# Print the top few observations of a variable
head(dataframe$vector)
# Print the bottom few observations of a variable
tail(dataframe$vector)
# Print all observations of a variable
print(dataframe$vector)
# Access a summary of a variable or model
summary(dataframe$vector)
summary(model)
```

To learn the structure or format of your data:
```{r eval=FALSE}
# Access the type / storage mode of the data
typeof(dataframe$vector) 
# Access the structure of the data
str(dataframe) 
str(dataframe$vector) 
# Access the length of a vector (e.g. the number of observations)
length(dataframe$vector)
# Access the attributes and metadata of an object
attributes(dataframe) 
attributes(dataframe$vector) 
```

All of the examples in this guide will use the [2016 CCES](https://doi.org/10.7910/DVN/GDF6Z0) loaded in `.dta` format. To replicate, 

```{r}
library(haven)
cces16 <- read_dta("../../Data/2016 CCES/CCES16_Common_OUTPUT_Feb2018_VV.dta")
```

# Manipulating Data
## Recoding Data
To recode values in base R:

```{r}
#Create a new vector to work with
cces16$ban_ar <- cces16$CC16_330d
#Attitudes toward gun control
#Call all values for "oppose" and replace with zero
cces16$ban_ar[cces16$ban_ar==2] <- 0
#Call all values for skipped / not asked and replace with missing
cces16$ban_ar[cces16$ban_ar==8] <- NA
cces16$ban_ar[cces16$ban_ar==9] <- NA
```

To collapse categories of an ordinal variable in base R:
```{r}
#Create a new vector to work with
cces16$pid <- cces16$pid7
#Recode values to missing
cces16$pid[cces16$pid==98] <- NA
cces16$pid[cces16$pid==99] <- NA
cces16$pid[cces16$pid==8] <- NA
#Collapse the categories
cces16$pid[cces16$pid==2] <- 1
cces16$pid[cces16$pid==3] <- 1
cces16$pid[cces16$pid==4] <- 2
cces16$pid[cces16$pid==5] <- 3
cces16$pid[cces16$pid==6] <- 3
cces16$pid[cces16$pid==7] <- 3
```


To cut a continuous variable into an ordinal one in base R:
```{r}
# Create the age variable
cces16$age <- 2016 - cces16$birthyr

cces16$agecats <- cut(cces16$age,
                      breaks=c(-Inf, 35, 50, Inf),
                      labels=c("35 and Under","36 to 50","Over 50"))
```

To flip the direction of coding in base R:
```{r}
#Flip the coding
cces16$pid_reverse <- 4 - cces16$pid
```

To apply labels to factor levels of a recoded variable:
```{r}
cces16$pid_reverse <- factor(cces16$pid_reverse,
                             levels = c(1,2,3),
                             labels = c("Republican", "Independent",
                             "Democrat"))
```


## Merging Data
To merge data that have one common vector with the same name:


```{r}
#Load the supplemental data
cces16s <-  read.csv("../../Data/2016 CCES Supplementary/CC16_Candidates_By_Race.csv")
#Merge the primary and supplemental data by respondent state
cces16c <- merge(cces16, cces16s)
```


To merge data that have a common vector with different names:
```{r}
#Merge by state with different column names
rm(cces16c)
cces16c <- merge(cces16, cces16s,
                 by.x = "inputstate", 
                 by.y = "inputstate")
```


#Descriptive Statistics
## Summary Statistics
To summarize a variable:

```{r}
summary(cces16$age)
```

To call specific summary statistics:
```{r}
#Mean
mean(cces16$age)
```
```{r}
#Standard deviation
sd(cces16$age)
```
```{r}
#Minimum
min(cces16$age)
```
```{r}
#Maximum
max(cces16$age)
```
```{r}
#Range
range(cces16$age)
```
```{r}
#Quantiles
quantile(cces16$age)
```
## Tabulations & Cross Tabulations

To tabulate a variable:
```{r}
# Tabulate PID
prop.table(table(cces16$pid_reverse))
```

To cross-tabulate two variables:
```{r}
# Tabulate PID by age categories
pidxage <- table(cces16$pid_reverse, cces16$agecats)
pidxage
prop.table(pidxage,2)
```

## Summary Statistics by Group
To summarize variables by group in base R:

```{r}
#Create a dataframe of the variables of interest
subgroup_vars <- c("age","pid")
subgroup_matrix <- as.matrix(cces16[subgroup_vars])
subgroup_df <- as.data.frame(subgroup_matrix)
aggregate(subgroup_df$age, list(subgroup_df$pid), mean)
```

To summarize variables by group with `dplyr`:

```{r}
library(dplyr)
subgroup_means <- subgroup_df %>%
  group_by(pid) %>%
  summarise(mean = mean(age))
subgroup_means
```

To visualize subgroup means with `ggplot`:

```{r}
library(ggplot2)
pid_labels <- c("Democrat", "Independent", "Republican")
sg_bp <- ggplot(subgroup_means, aes(y=mean, x=pid)) + 
  geom_bar(fill="lightskyblue",stat="identity") +
  #xlab("Partisanship") +
  ylab("Mean Age") +
  scale_x_discrete(name = "Partisanship",
                   limits=pid_labels)
sg_bp <- sg_bp + coord_cartesian(ylim=c(45,55))
sg_bp
```


## Distributions
To plot a histogram of a variable's distribution using `ggplot2`:

```{r}
#Histogram for age
library(ggplot2)
age_matrix <- as.matrix(cces16$age)
age_df <- as.data.frame(age_matrix)
ggplot(data = age_df, aes(x=age_df)) +
  geom_histogram(binwidth = 1) + 
  xlab("Age of Respondent") +
  ylab("Frequency")
```

```{r}
#Density plot
ggplot(data = age_df, aes(x=age_df)) +
  geom_density(fill="lightblue") +
  xlab("Age of Respondent") +
  ylab("Density")
```


# Modeling
## OLS Regression

```{r}
# Prepare Obama approval as the DV
cces16$oa <- cces16$CC16_320a
cces16$oa[cces16$oa=="5"] <- NA
cces16$oa[cces16$oa=="8"] <- NA
cces16$oa <- 5 - cces16$oa
# Fit the OLS model
olsfit <- lm(oa ~ pid + gender + age,
             data = cces16)
# Print the model summary
summary(olsfit)
```

## Logistic Regression
```{r}
## Logistic Regression
# Prepare preference for AR ban as DV
cces16$ban_ar <- cces16$CC16_330d
cces16$ban_ar[cces16$ban_ar==2] <- 0
cces16$ban_ar[cces16$ban_ar==8] <- NA
cces16$ban_ar[cces16$ban_ar==9] <- NA
# Fit the logit model
lfit <- glm(ban_ar ~ pid + agecats + gender,
           data=cces16, family = binomial())
# Print the model summary
summary(lfit)
```

## Ordinal Logistic Regression

```{r}
## Ordinal Logistic Regression
# Convert Obama approval to a factor
cces16$oaf <- factor(cces16$oa)
library(MASS)
# Fit the ordinal logit model
olfit <- polr(oaf ~ pid + gender + age,
              data = cces16, Hess = TRUE)
# Print the model summary
summary(olfit)
```

## Multinomial Logistic Regression
```{r}
## Multinomial Logistic Regression
## Create a 2012 vote choice factor variable
cces16$vc <- cces16$CC16_326
cces16$vc[cces16$vc=="4"] <- NA
cces16$vc[cces16$vc=="5"] <- NA
cces16$vc <- factor(cces16$vc,
                    levels = c(1,2,3),
                    labels = c("Obama", "Romney","Other"))

library(nnet)
# Fit the multinomial logit model
mlfit <- multinom(vc ~ pid + agecats + gender,
                  data=cces16)
# weights:  27 (16 variable)
```

```{r}
# Print the model summary
summary(mlfit)
```

# `survey` Package

All examples in this section require the `survey` package:
```{r}
library(survey)
```

## Weighted Tabulations
```{r}
## WITHOUT WEIGHTING 
# Tabulate PID
prop.table(table(cces16$pid_reverse))
```
```{r}
## WITHOUT WEIGHTING 
# Tabulate PID by age categories
pidxage <- table(cces16$pid_reverse, cces16$agecats)
pidxage
prop.table(pidxage,2)
```

```{r}
## WITH WEIGHTING
# Create a survey design dataframe
svy.cces16 <- svydesign(ids = ~1, 
                        data = cces16, 
                        weights = cces16$commonweight_vv)
# Weighted tabulation of PID
prop.table(svytable(~cces16$pid_reverse, design = svy.cces16))
```
```{r}
## WITH WEIGHTING
# Weighted tabulation of PID by age categories
prop.table(svytable(~cces16$pid_reverse+cces16$agecats, 
                    design = svy.cces16),2)
```

## Weighted Models
```{r}
## WITHOUT WEIGHTING 
# Run a logit regression for attitudes toward gun control
fit <- glm(ban_ar ~ pid_reverse + agecats + gender,
           data=cces16, family = binomial())
summary(fit)
```

```{r}
## WITH WEIGHTING
wfit <- svyglm(ban_ar ~ pid_reverse + agecats + gender,
               design=svy.cces16, family = binomial())
summary(wfit)
```


# Creating Post-Stratification Weights
In progress.



















